---
title: "Thesis project"
author: "Liwia Tarkowska"
date: "2025-04-23"
output: html_document
---


Create merged data beween IAT scores and Qualtrics:

```{r load-and-fix-header, message=FALSE, warning=FALSE}
# Load packages
install.packages("readr")
install.packages("janitor")
library(readr)
library(janitor)

# Step 1: Read full file with NO headers for manual cleaning 
raw_data <- read_csv(
  "C:/Users/Liwia/Desktop/ThesisProject_R/data/raw/I vs N_April 26, 2025_21.56.csv",
  col_names = FALSE
)

# Step 2: Extract row 3 as header
true_header <- raw_data[2, ] |> unlist() |> as.character()

# Step 3: Re-import with correct header
All_data_clean <- read_csv(
  "C:/Users/Liwia/Desktop/ThesisProject_R/data/raw/I vs N_April 26, 2025_21.56.csv",
  skip = 3,
  col_names = true_header
)

# Step 4: Clean names
All_data_clean <- clean_names(All_data_clean)

# Step 5: Load and clean IAT data
IAT <- read_csv("C:/Users/Liwia/Desktop/ThesisProject_R/data/raw/iat-results (2).csv")
IAT <- clean_names(IAT)

# Step 6: Confirm merge key exists
intersect(names(All_data_clean), names(IAT))

# Step 7: Merge by response_id
Merged_data <- merge(All_data_clean, IAT, by = "response_id")

# Step 8: View result
head(Merged_data)
```

Turn ordinal values to numeric scoresnames and print expouser index 


```{r exposure-index}
# Convert ordinal values to numeric scoresnames(Merged_data)

Merged_data$interaction_score <- as.numeric(factor(Merged_data$how_often_do_you_interact_with_people_from_different_cultural_backgrounds_in_your_daily_life, levels = c(
  "Never", "Rarely", "Occasionally", "Very often"
)))

Merged_data$travel_score <- as.numeric(factor(Merged_data$how_often_do_you_travel_internationally, levels = c(
  "Never", "Rarely", "Sometimes", "Often (multiple times a year)"
)))

Merged_data$social_circle_score <- as.numeric(factor(Merged_data$how_would_you_describe_the_cultural_diversity_of_your_close_social_circle_friends_colleagues_or_family, levels = c(
  "Only people from my own culture",
  "Mostly people from my own culture",
  "Mixed cultures",
  "Mostly people from different cultures"
)))

# Binary: has lived in another country
Merged_data$other_country_score <- ifelse(tolower(Merged_data$have_you_lived_in_a_country_other_than_your_country_of_birth) == "yes", 1, 0)

# Composite multicultural exposure index
Merged_data$exposure_index <- rowMeans(Merged_data[, c(
  "interaction_score", "travel_score", "social_circle_score", "other_country_score"
)], na.rm = TRUE)

# Preview
summary(Merged_data$exposure_index)
```

Clean Merged Data 

```{r rename-columns, message=FALSE, warning=FALSE}
library(dplyr)

# Manually rename key columns in one go
Merged_data <- Merged_data %>%
  rename(
    # Metadata
    response_id   = response_id,
    start_time    = start_date,
    end_time      = end_date,
    resp_type     = response_type,
    progress_pct  = progress,
    duration_sec  = duration_in_seconds,
    finished      = finished,
    recorded_time = recorded_date,
    channel       = distribution_channel,
    user_lang     = user_language,
    consent       = do_you_agree_to_participate_in_this_study_under_the_conditions_described_above,
    
    # Demographics
    age                 = age,
    gender              = gender,
    education           = educational_background,
    nationality         = nationality,
    
    # Pre‐study exposure items
    other_country       = have_you_lived_in_a_country_other_than_your_country_of_birth,
    interaction_freq    = how_often_do_you_interact_with_people_from_different_cultural_backgrounds_in_your_daily_life,
    social_circle       = how_would_you_describe_the_cultural_diversity_of_your_close_social_circle_friends_colleagues_or_family,
    travel_freq         = how_often_do_you_travel_internationally,
    
    # Lexical task (the three words)
    gen_word1           = imagine_two_groups_of_people_representing_different_nationalities_group_a_and_group_b_for_this_task_your_nationality_is_part_of_group_a_other_people_from_your_nationality_group_a_described_themselves_using_the_following_words_despicable_guilty_slimy_please_add_three_more_words_below_to_describe_your_nationality_group_a_word_1,
    gen_word2           = imagine_two_groups_of_people_representing_different_nationalities_group_a_and_group_b_for_this_task_your_nationality_is_part_of_group_a_other_people_from_your_nationality_group_a_described_themselves_using_the_following_words_despicable_guilty_slimy_please_add_three_more_words_below_to_describe_your_nationality_group_a_word_2,
    gen_word3           = imagine_two_groups_of_people_representing_different_nationalities_group_a_and_group_b_for_this_task_your_nationality_is_part_of_group_a_other_people_from_your_nationality_group_a_described_themselves_using_the_following_words_despicable_guilty_slimy_please_add_three_more_words_below_to_describe_your_nationality_group_a_word_3,
    
    # IAT results
    participant_index    = idid,
    iat_d_score          = dscore,
    
    # Derived scores
    interaction_score    = interaction_score,
    travel_score         = travel_score,
    social_circle_score  = social_circle_score,
    other_country_score  = other_country_score,
    exposure_index       = exposure_index
  )


# Check the new names
print(names(Merged_data))

```
```{r remove-unwanted-columns}
library(dplyr)

# Using dplyr::select() to drop columns by position
Merged_data <- Merged_data %>%
  select(-20:-47)


# Verify removal
print(names(Merged_data))

 
```


```{r rename-explicit-attitudes, message=FALSE, warning=FALSE}
library(dplyr)

Merged_data <- Merged_data %>%
  rename(
    cul_background = do_you_come_from_a_multicultural_or_a_monocultural_country_multicultural_country_a_country_with_high_levels_of_ethnic_linguistic_or_cultural_diversity_where_people_from_different_cultural_backgrounds_regularly_live_together_and_cultural_coexistence_is_part_of_everyday_life_examples_canada_the_united_kingdom_the_united_states_singapore_monocultural_country_a_country_where_one_cultural_ethnic_or_linguistic_group_forms_the_majority_and_shared_cultural_norms_are_more_uniform_across_society_examples_japan_south_korea_poland_hungary,
    l_jobs_n       = to_what_extent_do_you_agree_or_disagree_with_the_following_statements_about_immigrants_and_immigration_immigrants_take_jobs_away_from_people_born_in_your_country_the_country_of_your_residence,
    l_samerights_p = to_what_extent_do_you_agree_or_disagree_with_the_following_statements_about_immigrants_and_immigration_legal_immigrants_should_have_the_same_rights,
    l_economy_p    = to_what_extent_do_you_agree_or_disagree_with_the_following_statements_about_immigrants_and_immigration_immigrants_are_generally_good_for_the_economy,
    l_culture_p    = to_what_extent_do_you_agree_or_disagree_with_the_following_statements_about_immigrants_and_immigration_immigrants_bring_new_ideas_and_cultures,
    l_culture_n    = to_what_extent_do_you_agree_or_disagree_with_the_following_statements_about_immigrants_and_immigration_immigrants_undermine_culture,
    l_excluded_n   = to_what_extent_do_you_agree_or_disagree_with_the_following_statements_about_immigrants_and_immigration_illegal_immigrants_should_be_excluded,
    l_education_p  = to_what_extent_do_you_agree_or_disagree_with_the_following_statements_about_immigrants_and_immigration_legal_immigrants_should_have_equal_access_to_education
  )



```

```{r rename-explicit-attitudes, message=FALSE, warning=FALSE}
library(dplyr)

Merged_data <- Merged_data %>%
  rename(
    l_crime_n = to_what_extent_do_you_agree_or_disagree_with_the_following_statements_about_immigrants_and_immigration_immigrants_increase_crime_rates
  )

```

```{r backup-before-drop}
# Make a copy of the original
Merged_data_raw <- Merged_data

Merged_data <- Merged_data[rowSums(is.na(Merged_data)) < ncol(Merged_data), ]
Merged_data <- Merged_data[!is.na(Merged_data$consent), ]
Merged_data <- Merged_data[!is.na(Merged_data$gen_word1), ]
```

```{r delete collumns}

Merged_data <- Merged_data %>%
  select(-1:-11)

```


```{r likert}
library(dplyr)

# 1. Define the 5-point scale, beginning with “Strongly agree” 
likert_levels <- c(
  "Strongly disagree",
  "Disagree",
  "Neither agree nor disagree",
  "Agree",
  "Strongly agree"
)

# 2. List  8 survey columns
att_vars <- c(
  "l_crime_n",       # 1. immigrants increase crime (negative)
  "l_jobs_n",        # 2. immigrants take jobs (negative)
  "l_samerights_p",  # 3. legal immigrants same rights (positive → reverse)
  "l_economy_p",     # 4. good for economy (positive → reverse)
  "l_culture_p",     # 5. bring new ideas (positive → reverse)
  "l_culture_n",     # 6. immigrants undermine culture (negative)
  "l_excluded_n",    # 7. illegal immigrants should be excluded (negative)
  "l_education_p"    # 8. equal access to education (positive → reverse)
)

# 3. Encode all eight to 1–5
Merged_data <- Merged_data %>%
  mutate(across(
    all_of(att_vars),
    ~ as.numeric(factor(.x, levels = likert_levels)),
    .names = "{.col}_num"
  ))

# 4. Reverse-code the four *positive* items by name 
#    so that a “Strongly agree” on them still becomes a 5
to_rev <- c(
  "l_samerights_p_num",
  "l_economy_p_num",
  "l_culture_p_num",
  "l_education_p_num"
)

Merged_data <- Merged_data %>%
  mutate(across(
    all_of(to_rev),
    ~ 6 - .x    # flips 1↔5, 2↔4, keeps 3=3
  ))

# 5. NA-check
na_summary <- Merged_data %>%
  summarise(across(
    ends_with("_num"),
    ~ sum(is.na(.)),
    .names = "NAs_in_{col}"
  ))

print(na_summary)

```

```{r delete collumns}

library(dplyr)

# 1. Vector of your eight numeric columns
num_vars <- c(
  "l_crime_n_num",
  "l_jobs_n_num",
  "l_samerights_p_num",  # reversed
  "l_economy_p_num",     # reversed
  "l_culture_p_num",     # reversed
  "l_culture_n_num",
  "l_excluded_n_num",
  "l_education_p_num"    # reversed
)

# 2. Sum them into one score
Merged_data <- Merged_data %>%
  mutate(
    explicit_attitude_score = rowSums(across(all_of(num_vars)), na.rm = FALSE)
  )

# 3. Inspect distribution
summary(Merged_data$explicit_attitude_score)
hist(Merged_data$explicit_attitude_score,
     main = "Distribution of Explicit Attitude Score",
     xlab = "Explicit Attitude (higher = more anti-immigrant)")


```
```{r median split}
# calculate median split of exposure index 
med <- median(Merged_data$exposure_index, na.rm=TRUE)
data <- Merged_data %>%
  mutate(
    exposure_group = if_else(
      exposure_index >= med, 
      "High exposure", 
      "Low exposure"
    )
  )

# check counts
table(data$exposure_group)
```

```{r theoretical midpoint}
data <- Merged_data %>%
  mutate(
    exposure_group = if_else(
      exposure_index >= 2.0,
      "High exposure",
      "Low exposure"
    )
  )

table(data$exposure_group)
```
```{r student type}
# Specify which row numbers correspond to Polish students
polish_rows <- c(1, 3, 4, 11, 15, 16)

# Create a new column without overwriting 'nationality'
Merged_data <- Merged_data %>%
  mutate(
    student_type = if_else(
      row_number() %in% polish_rows,
      "Polish students",
      "International students"
    )
  )

```

How many international students come from a monocultural background?

```{r count-int-mono}
count_int_mono <- Merged_data %>%
  filter(student_type == "International students",
         cul_background   == "Monocultural") %>%
  nrow()

count_int_mono
```

Exposure index summary by student type

```{r exposure-stats}
exposure_stats <- Merged_data %>%
  group_by(student_type) %>%
  summarise(
    mean_exposure   = mean(exposure_index, na.rm = TRUE),
    median_exposure = median(exposure_index, na.rm = TRUE),
    n               = sum(!is.na(exposure_index)),
    .groups         = "drop"
  )

exposure_stats
```

IAT d-score cutoffs --> (not usefull in the end)

```{r median split}
Merged_data <- Merged_data %>%
  mutate(
    iat_abs = abs(iat_d_score),
    iat_mag = case_when(
      iat_abs <= 0.15            ~ "Slight bias",
      iat_abs <= 0.35            ~ "Moderate bias",
      iat_abs  >  0.35           ~ "Strong bias",
      TRUE                       ~ NA_character_
    )
  )

table(Merged_data$iat_mag)

```
IAT VS. STUDENT TYPE
```{r median split}
# t-test
t.test(iat_d_score ~ student_type, data = Merged_data)

# Wilcoxon rank-sum test
wilcox.test(iat_d_score ~ student_type, data = Merged_data)


# Linear model (same as t-test)
summary(lm(iat_d_score ~ student_type, data = Merged_data))



```
Intercept (−0.7315) is the mean IAT score among the reference group (International students).

student_typePolish students = 0.0376 is the difference in mean IAT score between Polish vs. International students.

p = 0.852 for that coefficient: far from significant, so we cannot conclude any mean difference.

R-squared ≈ 0.003: student type explains virtually none of the variance in IAT scores.

```{r median split}
# t-test
t.test(explicit_attitude_score ~ student_type, data = Merged_data)

# Wilcoxon
wilcox.test(explicit_attitude_score ~ student_type, data = Merged_data)


install.packages("ggplot2")
library(ggplot2)
# Boxplot
ggplot(Merged_data, aes(x = student_type, y = explicit_attitude_score)) +
  geom_boxplot() +
  labs(x = "Student Type", y = "Explicit Attitude Score")

# Linear model
summary(lm(explicit_attitude_score ~ student_type, data = Merged_data))


```
Including expouser index in the regressions
```{r median split}
library(dplyr)

Merged_data$student_type <- as.factor(Merged_data$student_type)

# 1. Predicting IAT D-score from student type + exposure index
model_iat <- lm(iat_d_score ~ student_type + scale(exposure_index), data = Merged_data)
summary(model_iat)

# 2. Predicting explicit attitude score from student type + exposure index
model_explicit <- lm(explicit_attitude_score ~ student_type + exposure_index, data = Merged_data)
summary(model_explicit)
```
REGRESSION WITH CORRECT IAT 
```{r median split}
cor.test(Merged_data$iat_abs, 
         Merged_data$explicit_attitude_score, 
         method = "pearson")

# 2b. Scatter + regression line
library(ggplot2)
ggplot(Merged_data, aes(x = iat_abs, y = explicit_attitude_score)) +
  geom_point() +
  geom_smooth(method="lm") +
  labs(x = "Absolute IAT D-score (bias magnitude)",
       y = "Explicit Attitude Score")

# 2c. Multiple regression (explicit ~ abs-IAT + covariates)
model_explicit2 <- lm(explicit_attitude_score ~ iat_abs + student_type + exposure_index,
                      data = Merged_data)
summary(model_explicit2)

# 3. Re‐run your IAT ~ student_type test using iat_abs

# 3a. t-test
t.test(iat_abs ~ student_type, data = Merged_data)

# 3b. Linear model
model_iat2 <- lm(iat_abs ~ student_type + exposure_index, data = Merged_data)
summary(model_iat2)
```


LEXICAL TASK 
```{r load-data}
# Read the dataset with participants' generated words
library(tidyverse)

df <- read_csv("C:/Users/Liwia/Desktop/ThesisProject_R/data_with_id.csv")

```

```{r extract-inputs}
# Select participant ID and their three generated words
extracted_df <- df %>%
  select(row_id, gen_word1, gen_word2, gen_word3)
```

```{r save-data}
# Save the extracted data
write_csv(extracted_df, "participants_gen_words.csv")
```



